---
title: "Data Mining Final Project"
author: "Pia Chouaifaty"
date: "12/14/2020"
output: html_document
---

Libraries
```{r}
library(leaps)
library(ggbiplot)
library(pls)

```


## Summary of Attributes

Matrix column entries (attributes):
name - ASCII subject name and recording number
MDVP:Fo(Hz) - Average vocal fundamental frequency
MDVP:Fhi(Hz) - Maximum vocal fundamental frequency
MDVP:Flo(Hz) - Minimum vocal fundamental frequency
MDVP:Jitter(%),MDVP:Jitter(Abs),MDVP:RAP,MDVP:PPQ,Jitter:DDP - Several 
measures of variation in fundamental frequency
MDVP:Shimmer,MDVP:Shimmer(dB),Shimmer:APQ3,Shimmer:APQ5,MDVP:APQ,Shimmer:DDA - Several measures of variation in amplitude
NHR,HNR - Two measures of ratio of noise to tonal components in the voice
status - Health status of the subject (one) - Parkinson's, (zero) - healthy
RPDE,D2 - Two nonlinear dynamical complexity measures
DFA - Signal fractal scaling exponent
spread1,spread2,PPE - Three nonlinear measures of fundamental frequency variation

## Reading in the File

```{r}
pd = read.csv("/Users/piachouaifaty/parkinson.csv")
```

```{r}
length(which(pd$status==0))
```
48 measurements for healthy patients  

```{r}
length(which(pd$status==1))
```
147 measurements for PD patients  

## Averaging Repeated Measurements
```{r}
pdavg = pd
```

Removing the recording number from the names to be able to match and average them
```{r}
#substitutes the characters after the last underscore with a "", so removes them
pdavg$name = sub("_[^_]+$", "", pdavg$name)
```

```{r}
head(pdavg)
```

Averaging all the rows using aggregate by name (after removing the recording number, we are left with the individual ids)
```{r}
pdavg=aggregate(pdavg[,2:24], list(pdavg$name), mean)
head(pdavg)
```

```{r}
length(which(pdavg$status==0))
length(which(pdavg$status==1))
```

8 healthy patients and 24 patients with PD  

## Analysis on Full Dataset (All Measurements)

```{r}
pd$status=as.factor(pd$status)
```

```{r}
summary(pd)
```

### Plotting for Visualization

```{r}
pairs(pd[,-1], col=pd$status)
#red=parkinson's
```

```{r}
plot(pd$PPE, pd$spread1, col=pd$status)
```

```{r}
plot(pd$PPE, pd$spread1, col=pd$status)
```


```{r}
pd2=pd[,-1]
rownames(pd2) = pd[,1]
pd2=pd2[,-17]
par(mfrow=c(3,3))
for (i in colnames(pd2))
  {hist(pd2[,i], main=i)}

```

```{r}
par(mfrow=c(1,3))
for (i in colnames(pd2))
  {boxplot(pd2[,i], main=i)}
```

```{r}
outl=boxplot.stats(pd2[,"MDVP.Flo.Hz."])$out
ind=which(pd2[,"MDVP.Flo.Hz."] %in% c(outl))
pd2[c(ind),]
```

```{r}
pd[c(ind),]
```

I check some of the outliers in the MDVP.Flo.Hz. column.  
5 of them come from the same individual, and the number of "outliers" is very high. Also, the data is skewed in general, so I don't remove any of the outliers. Since the majority of measurements are diseased, and the outliers in one of the columns are all control, I think all the "outliers" are actually just control measurements.  
I check a few more.

```{r}
outl=boxplot.stats(pd2[,"MDVP.Fhi.Hz."])$out
ind=which(pd2[,"MDVP.Fhi.Hz."] %in% c(outl))
pd[c(ind),]
```
In this case many of the outlier measurements come from the same individual.  

```{r}
outl=boxplot.stats(pd2[,"MDVP.Jitter..."])$out
ind=which(pd2[,"MDVP.Jitter..."] %in% c(outl))
pd[c(ind),]
```
In this case, outliers come from the same individual.


```{r}
outl=boxplot.stats(pd2[,"MDVP.RAP"])$out
ind=which(pd2[,"MDVP.RAP"] %in% c(outl))
pd[c(ind),]
```
Here, two individuals are "outliers."  
Dealing with outliers by either removing them or replacing them with the mean/median seems unnecessary - the data is just skewed in nature. Removing outliers would make the analysis biased.  
So I decide to keep them.
Since the data is not normally distributed (many predictors are skewed): will have to normalize it when working with LDA and QDA.

### Unsupervised Learning Methods for Visualization

```{r}
pca_pd=prcomp(pd2, scale=TRUE)
```

```{r}
str(pca_pd)
```

```{r}
#library(devtools)
#install_github("vqv/ggbiplot")
#library(ggbiplot)
```

```{r}
ggbiplot(pca_pd)
```

```{r}
biplot(pca_pd, scale=0)
```


```{r}
ggbiplot(pca_pd, labels=rownames(pd2))
```

```{r}
disease_stat=ifelse(pd$status=="1", "PD", "HLT")
ggbiplot(pca_pd,ellipse=TRUE, groups=disease_stat)
```

The PD measurements seem to be clustered to the bottom

```{r}
pca_pd$rotation
```

Proprotion of Variance Explained by PC1

```{r}
pr.var=pca_pd$sdev^2
pve=pr.var/sum(pr.var)
par(mfrow=c(1,2))
plot(pve, xlab="Principal Component", ylab="Proportion of Variance Explained ", ylim=c(0,1),type="b")
plot(cumsum(pve), xlab="Principal Component ", ylab=" Cumulative Proportion of Variance Explained ", ylim=c(0,1), type="b")
```

PC1 explains 58.9% of the variance.

#### Clustering

Clustering with K=2
```{r}
km.pd2=kmeans(pd2,2,nstart=20)
```

```{r}
head(km.pd2$cluster)
```


```{r}
head(pd$status)
s=ifelse(pd$status==1, 1, 2) #so the colors match
head(s) 
```

```{r}
par(mfrow=c(1,2))
for (i in colnames(pd2))
{plot(pd2[,i], col=(km.pd2$cluster), main="K-Means Clustering; K=2", xlab="", ylab=i, pch=20, cex=2)
plot(pd2[,i], col=(s), main="By Status; black=PD", xlab="", ylab=i, pch=20, cex=2)}
```

#### Normalizing the Data

```{r}
pd2=cbind(pd2, pd$status)
```

```{r}
colnames(pd2)=c("MDVP.Fo.Hz.",      "MDVP.Fhi.Hz." ,    "MDVP.Flo.Hz."  ,   "MDVP.Jitter..." ,  "MDVP.Jitter.Abs." ,"MDVP.RAP"     ,   "MDVP.PPQ"    ,     "Jitter.DDP"  ,     "MDVP.Shimmer"   ,  "MDVP.Shimmer.dB." ,"Shimmer.APQ3"   ,  "Shimmer.APQ5" ,   
 "MDVP.APQ"  ,       "Shimmer.DDA"  ,    "NHR" ,             "HNR"    ,          "RPDE"    ,         "DFA"     ,        
"spread1"     ,     "spread2"      ,    "D2"        ,       "PPE"      ,        "status")
```

The data is skewed, so in order to normalize it, I log transform all the values.
The values for spread1 are negative, so I add a constant to all the values.

```{r}
offset=min(pd2[,"spread1"])
offset=-offset
offset
```


```{r}
pd2norm=log(offset+1+pd2[1:22])
pd2norm=cbind(pd2norm, pd2[,"status"])
pd2norm
```

```{r}
colnames(pd2norm)=c("MDVP.Fo.Hz.",      "MDVP.Fhi.Hz." ,    "MDVP.Flo.Hz."  ,   "MDVP.Jitter..." ,  "MDVP.Jitter.Abs." ,"MDVP.RAP"     ,   "MDVP.PPQ"    ,     "Jitter.DDP"  ,     "MDVP.Shimmer"   ,  "MDVP.Shimmer.dB." ,"Shimmer.APQ3"   ,  "Shimmer.APQ5" ,   
 "MDVP.APQ"  ,       "Shimmer.DDA"  ,    "NHR" ,             "HNR"    ,          "RPDE"    ,         "DFA"     ,        
"spread1"     ,     "spread2"      ,    "D2"        ,       "PPE"      ,        "status")
```

Plotting Again

```{r}
par(mfrow=c(3,3))
for (i in colnames(pd2norm[1:22]))
  {hist(pd2norm[,i], main=i)}
```

```{r}
par(mfrow=c(1,3))
for (i in colnames(pd2norm[1:22]))
  {boxplot(pd2norm[,i], main=i)}
```

PCA Again

```{r}
pca_pd=prcomp(pd2norm[1:22], scale=TRUE)
```

```{r}
str(pca_pd)
```

```{r}
#library(devtools)
#install_github("vqv/ggbiplot")
#library(ggbiplot)
```

```{r}
ggbiplot(pca_pd)
```

```{r}
biplot(pca_pd, scale=0)
```


```{r}
ggbiplot(pca_pd, labels=rownames(pd2))
```

```{r}
disease_stat=ifelse(pd$status=="1", "PD", "HLT")
ggbiplot(pca_pd,ellipse=TRUE, groups=disease_stat)
```

The PD measurements seem to be clustered to the bottom

```{r}
#pca_pd$rotation
```


#### Training and Test Sets

Since there are only 8 healthy individuals (around 84 measurements for healthy vs 147 for diseased), I make sure to include a sufficient number in the training and test sets, so I split them into healthy vs diseased, sample from each, and then combine them into training and test sets.

```{r}
pd_dis_idx=which(pd2$status==1) #indeces of PD individuals
pd_health_idx=which(pd2$status==0) #indeces of healthy individuals
#pd_health_idx
#pd_dis_idx
set.seed(4706)

#nrow(pd_dis)
#nrow(pd_health)

#73 diseased test
#24 healthy test
#74 diseases train
#24 healthy train

test_health=sample(pd_health_idx, 24, replace = FALSE) #sample from indeces
#test_health

test_dis=sample(pd_dis_idx, 73, replace = FALSE) #sample from indeces
#test_dis

test=append(test_dis, test_health) #test set

tot=1:nrow(pd2)

train=tot[-test] #training set

y.train=pd2[c(train), "status"]
y.test=pd2[c(test), "status"]

```

### Logistic Regression

I fit a logistic regression model on the training data

```{r}
logist_reg_fit=glm(status~., data=pd2, family=binomial, subset=train)

summary(logist_reg_fit)

#"Coefficients"
#coef(logist_reg_fit)

#"P-Values"
#summary(logist_reg_fit)$coef[ ,4]
```

At first glance, the p-values are all very large, except for DFA, which is smaller but still pretty large.

```{r}
contrasts(pd2[,"status"])
```

```{r}
glm_probs = predict(logist_reg_fit, pd2[test,], type="response")
head(glm_probs)
glm_pred=ifelse(glm_probs>0.5, 1, 0)
head(glm_pred)
```

```{r}
length(glm_pred)
length(y.test)
table(glm_pred, y.test)
```

Classification Error
```{r}
mean(glm_pred!=y.test)
```
Accuracy
```{r}
mean(glm_pred==y.test)
```


### LDA
```{r fitting the model}
lda_fit = lda(status~., data=pd2[,-5], subset=train)
lda_fit
plot(lda_fit) #produces plots of the linear discriminants
```

```{r }
lda_pred = predict(lda_fit, pd2[test,])
names(lda_pred) #the names are  class, contains LDA’s predictions

lda_class = lda_pred$class #the class predictions of our fitted model on test data
table(lda_class, y.test) #predictions vs real values of test data
contrasts(pd2[,"status"]) #the one with value=1 is the one for which we are getting the posterior probability

"Accuracy"
mean(lda_class==y.test)
"Classification Error"
mean(lda_class!=y.test)

#FNR=(FN/total positives)x100
#FPR=(FP/total negatives)x100 = 1-specificity = Type I error
#overall error=(FP+FN)/total x100
#sensitivity = TPR = (TP/Total positives) x100

sum(lda_pred$posterior[,1]>=0.5) #observations with predicted proba >0.5
sum(lda_pred$posterior[,1]<0.5)

#MY OWN POSTERIOR PROBABILITY THRESHOLD
#sum(lda_pred$posterior[,1]>0.9) #observations with predicted proba >0.9
#my_threshlold_classes=ifelse(lda_pred$posterior[,1]>0.9, "yes", "no") #yes being the one with contrast=1
```

```{r}
#CV for LDA
k=10 #number of folds
Dataset=pd2

set.seed(1)
folds=sample(1:k, nrow(Dataset), replace=TRUE)
#folds
Dataset=cbind(Dataset, folds)
table(folds) #to see that they are balanced
error=vector()#error for each fold

for (j in 1:k) #j refers to folds
{
  lda_fit = lda(status~., data=Dataset[folds!=j,-5]) #this is the training data, those NOT in the current fold

  lda_pred = predict(lda_fit, Dataset[folds==j,]) #test data, current fold
  lda_class = lda_pred$class #the class predictions of our fitted model on test data
  error=append(error, mean(lda_class!=Dataset[folds==j, "status"])) #cv error
}

cv_error=mean(error)
cv_error
```

### QDA

```{r fitting the model}
qda_fit=qda(status~., data=pd2, subset = train)
qda_fit

qda_class=predict(qda_fit, pd2[test,])$class
table(qda_class, y.test)

"Accuracy"
mean(qda_class==y.test)

"Classification Error"
mean(qda_class!=y.test)
#FNR=(FN/total positives)x100
#FPR=(FP/total negatives)x100 = 1-specificity = Type I error
#overall error=(FP+FN)/total x100
#sensitivity = TPR = (TP/Total positives) x100
```

Using the validation set approach, LDA performs best

```{r}
#CV for QDA
k=10 #number of folds
Dataset=pd2

set.seed(4706)
folds=sample(1:k, nrow(Dataset), replace=TRUE)
#folds
Dataset=cbind(Dataset, folds)
table(folds) #to see that they are balanced
error=vector()#error for each fold

for (j in 1:k) #j refers to folds
{
  qda_fit = qda(status~., data=Dataset[folds!=j,-5]) #this is the training data, those NOT in the current fold

  qda_class=predict(qda_fit, Dataset[folds==j,])$class #test data, current fold
  error=append(error, mean(qda_class!=Dataset[folds==j, "status"])) #cv error
}

cv_error=mean(error)
cv_error
```


### Trees





#### Best Subset Selection

```{r}
regfit_full=regsubsets(status~., data=pd2, subset=train, nvmax=23) #nvmax=p
reg_full_sum=summary(regfit_full) #returns R^2, adjusted R^2, AIC, BIC, Cp
reg_full_sum
names(reg_full_sum)
names(regfit_full)
```

*Plotting*
```{r plotting RSS, BIC...}
par(mfrow=c(2,2))

#RSS
plot(reg_full_sum$rss, xlab="Number of variables", ylab="RSS", type ="l")
mnrss=which.min(reg_full_sum$rss)
#Plotting min RSS
points(mnrss, reg_full_sum$adjr2[mnrss], col="red", cex=2, pch=20)

#Adjusted R^2
plot(reg_full_sum$adjr2, xlab="Number of variables", ylab="Adjusted R^2", type ="l")
mxr2=which.max(reg_full_sum$adjr2)
#Plotting max adjusted R^2
points(mxr2, reg_full_sum$adjr2[mxr2], col="red", cex=2, pch=20)

#CP
plot(reg_full_sum$cp, xlab = "Number of variables", ylab="CP", type = "l")
mncp=which.min(reg_full_sum$cp)
points(mncp, reg_full_sum$cp[mncp], col="blue", cex=2, pch=20)

#BIC
plot(reg_full_sum$bic, xlab="Number of variables", ylab = "BIC", type="l")
mnbic=which.min(reg_full_sum$bic)
points(mnbic, reg_full_sum$bic[mnbic], col="green", pch=20, cex=2)
```

BIC selects for a smaller model (5 predictors) (heavier penalty)  
CP min chooses around 6.  
Adjusted R^2 max is around 8.  
As expected, RSS decreases as more predictors are added, but this is not indicative of test error rate.

*Plotting Selected Variables*
```{r plotting variables according to BIC, RSS...}
plot(regfit_full, scale="r2")
plot(regfit_full, scale="adjr2")
plot(regfit_full, scale="Cp")
plot(regfit_full, scale="bic")
```

*Coef of models*
```{r}
"model with highest adjusted R^2"
coef(regfit_full, mxr2) #model with highest adjusted R^2
"model with lowest CP"
coef(regfit_full, mncp) #model with lowest CP
"model with lowest BIC"
coef(regfit_full, mnbic) #model with lowest BIC
```

### Cross Validating Model Size

Test Matrix
```{r}
test.mat=model.matrix(status~.,data=pd2[test,]) #output of the test dataset
head(test.mat)
```

```{r}
k=10
p=22
set.seed(1)
folds=sample(1:k, nrow(pd2), replace=TRUE)
folds
table(folds) #to see that it's pretty much balanced
cv.errors=matrix(NA, k, p, dimnames = list(NULL, paste (1:p)))
#k rows for each fold
#p columns for the model complexities (predictors)
```

```{r}
for (j in 1:k) #j refers to folds, i refers to predictors
{
  best.fit=regsubsets(Salary~.,data=Hitters[folds!=j,], nvmax=p) #this is the training data, those NOT in the current fold
      for (i in 1:p)
      {
        pred=predict(best.fit, Hitters[folds==j,], id=i) #this is the test data, those in the current fold
        #id = i meaning for each subset size from 1:p
        #we only called predict, not predict.regsubsets because we wrote our function in a way that the predict function           recognizes -> we give it a regsubsets object
        cv.errors[j,i]=mean((Hitters$Salary[folds==j]-pred)^2) #compute mean squared error
      }
}

cv.errors
```


#### Forward Selection

```{r}
regfit_fwd = regsubsets(status~.,pd2[train,], nvmax=23, method = "forward")
summary(regfit_fwd)
reg_fwd_sum=summary(regfit_fwd) #returns R^2, adjusted R^2, AIC, BIC, Cp
reg_fwd_sum
names(reg_fwd_sum)
names(regfit_fwd)

```
*Plotting*
```{r}
par(mfrow=c(2,2))

#RSS
plot(reg_fwd_sum$rss, xlab="Number of variables", ylab="RSS", type ="l")
mnrss=which.min(reg_fwd_sum$rss)
#Plotting min RSS
points(mnrss, reg_fwd_sum$adjr2[mnrss], col="red", cex=2, pch=20)

#Adjusted R^2
plot(reg_fwd_sum$adjr2, xlab="Number of variables", ylab="Adjusted R^2", type ="l")
mxr2=which.max(reg_fwd_sum$adjr2)
#Plotting max adjusted R^2
points(mxr2, reg_fwd_sum$adjr2[mxr2], col="red", cex=2, pch=20)

#CP
plot(reg_fwd_sum$cp, xlab = "Number of variables", ylab="CP", type = "l")
mncp=which.min(reg_fwd_sum$cp)
points(mncp, reg_fwd_sum$cp[mncp], col="blue", cex=2, pch=20)

#BIC
plot(reg_fwd_sum$bic, xlab="Number of variables", ylab = "BIC", type="l")
mnbic=which.min(reg_fwd_sum$bic)
points(mnbic, reg_fwd_sum$bic[mnbic], col="green", pch=20, cex=2)
```

*Plotting Selected Variables*
```{r}
plot(regfit_fwd, scale="r2")
plot(regfit_fwd, scale="adjr2")
plot(regfit_fwd, scale="Cp")
plot(regfit_fwd, scale="bic")
```

*Coef of models*
```{r}
"model with highest adjusted R^2"
coef(regfit_fwd, mxr2) #model with highest adjusted R^2
"model with lowest CP"
coef(regfit_fwd, mncp) #model with lowest CP
"model with lowest BIC"
coef(regfit_fwd, mnbic) #model with lowest BIC
```

#### Backward Selection

```{r}
regfit_bwd = regsubsets(status~.,pd2[train,], nvmax=23, method = "backward")
summary(regfit_bwd)
reg_bwd_sum=summary(regfit_bwd) #returns R^2, adjusted R^2, AIC, BIC, Cp
reg_bwd_sum
names(reg_bwd_sum)
names(regfit_bwd)
```

