---
title: "Data Mining Final Project"
author: "Pia Chouaifaty"
date: "12/14/2020"
output: html_document
---

### Summary of Attributes

Matrix column entries (attributes):
name - ASCII subject name and recording number
MDVP:Fo(Hz) - Average vocal fundamental frequency
MDVP:Fhi(Hz) - Maximum vocal fundamental frequency
MDVP:Flo(Hz) - Minimum vocal fundamental frequency
MDVP:Jitter(%),MDVP:Jitter(Abs),MDVP:RAP,MDVP:PPQ,Jitter:DDP - Several 
measures of variation in fundamental frequency
MDVP:Shimmer,MDVP:Shimmer(dB),Shimmer:APQ3,Shimmer:APQ5,MDVP:APQ,Shimmer:DDA - Several measures of variation in amplitude
NHR,HNR - Two measures of ratio of noise to tonal components in the voice
status - Health status of the subject (one) - Parkinson's, (zero) - healthy
RPDE,D2 - Two nonlinear dynamical complexity measures
DFA - Signal fractal scaling exponent
spread1,spread2,PPE - Three nonlinear measures of fundamental frequency variation

### Reading in the File

```{r}
pd = read.csv("/Users/piachouaifaty/parkinson.csv")
```

```{r}
length(which(pd$status==0))
```
48 measurements for healthy patients  

```{r}
length(which(pd$status==1))
```
147 measurements for PD patients  

### Averaging Repeated Measurements
```{r}
pdavg = pd
```

Removing the recording number from the names to be able to match and average them
```{r}
#substitutes the characters after the last underscore with a "", so removes them
pdavg$name = sub("_[^_]+$", "", pdavg$name)
```

```{r}
pdavg
```

Averaging all the rows using aggregate by name (after removing the recording number, we are left with the individual ids)
```{r}
pdavg=aggregate(pdavg[,2:24], list(pdavg$name), mean)
pdavg
```

```{r}
length(which(pdavg$status==0))
length(which(pdavg$status==1))
```

8 healthy patients and 24 patients with PD  

### Analysis on Full Dataset (All Measurements)

```{r}
pd$status=as.factor(pd$status)
```

```{r}
summary(pd)
```

#### Plotting for Visualization

```{r}
pairs(pd[,-1], col=pd$status)
#red=parkinson's
```

```{r}
plot(pd$PPE, pd$spread1, col=pd$status)
```

```{r}
plot(pd$PPE, pd$spread1, col=pd$status)
```

#### Unsupervised Learning Methods for Visualization

```{r}
pd2=pd[,-1]
rownames(pd2) = pd[,1]
pd2=pd2[,-17]
pr.out=prcomp(pd, scale=TRUE)
```

