---
title: "Scripts"
author: "Pia Chouaifaty"
date: "11/4/2020"
output: html_document
---

### Reading Files

*CSV*
```{r csv}
college = read.csv("/Users/piachouaifaty/ISLR_Data/College.csv")
#read.csv2 in case the separator is a semicolon instead of a comma
#read.delim 
#na.strings = "EMPTY" to set NA values as "EMPTY" instead
```

*TXT*
```{r txt}
Auto=read.table("/Users/piachouaifaty/ISLR_Data/Auto.data", header=T, na.strings = "?")
```

*EXCEL*
```{r excel}
library(xlsx)
tbl1 = read.xlsx("/Users/piachouaifaty/Book1.xlsx", sheetIndex = 1)
```

### Data Pre-Processing

*Checking Dimensions*
```{r dimensions} 
dim(dataset)
```

*Setting Rownames*
```{r setting rownames}
rownames(college)=college[,1]
```

*Deleting column*
```{r deleting columns}
college=college[,-1]
```

*Standardizing the data (usually for KNN)*
```{r standardizing for KNN}
standardizedX=scale(Dataset[,-qualitative_variable_index])
#sd=1, mean=0 for every predictor
```



#### Dealing with NAs

*Checking how many NAs*
```{r how many NAs}
length(which(is.na(dataset)))
```

*Removing NAs*
```{r removing NAs}
#(only if explicitly asked)
Dataset=na.omit(Dataset)
```

#### Statistics about Predictors

*Ranges*
```{r ranges for quantitative}
v=c("mpg", "cylinders", "displacement", "horsepower", "weight", "acceleration", "year")
# v is a vector of quantitative predictors
sapply(Dataset[,v], range)

#if they are already all quantitative
t(sapply(dataset, range))
```

*Means, SDs, Medians*

```{r means, sds, medians}
#where v is a vector of the quantitative predictor names

sapply(Dataset[,v], mean)
sapply(Dataset[,v], sd)
sapply(Dataset[,v], median)

#for individual predictor
mean(Dataset$predictor)
sd(Dataset$predictor)
median(Dataset$predictor)
```

*Min & Max*
```{r min & max}
#values for all predictors for observation with the max value in "predictor"
t(subset(Dataset, predictor == min(Dataset$predictor)))
t(subset(Dataset, predictor == max(Dataset$predictor)))
```

*Removing Observations*
```{r removing observations}
#(Range)
Dataset2 = Dataset[-(10:85),] 
```

*As factor*
```{r as factor for binary qualitative predictior}
#CAREFUL, if you declare as factor, CANNOT CALCULATE COR
dataset$predictor=as.factor(dataset$predictor)
```

*Creating new Predictor, Binning*
```{r binning - new predictor}
#1. New predictor "Elite"
Elite=rep("No", nrow(college))

#2. Condition for new predictor
Elite[college$Top10perc>50]="Yes"

#3. Transform to qualitative binary factor
Elite=as.factor(Elite)

#4. Add to dataset
college=cbind(college, Elite)
#Auto = data.frame(Auto, mpg01)

```

*Checking how many observations satisfy a condition*
```{r observations that satisfy a condition}

length((Dataset$predictor[Dataset$predictor>20]))
#or
dim(subset(Dataset, predictor == 1)) 
```

### Plotting

```{r output dimensions}
par(mfrow=c(1,1))
```

*Scatterplot*
```{r Scatterplot}
#PAIRED SCATTERPLOT
pairs(Smarket, col=Smarket$Direction) #[can only use quantitative & qualitative AFTER transforming to factor]
#col to color by category

#SCATTERPLOT OF 2 PREDICTORS
plot(college$Accept/college$Apps, college$S.F.Ratio) 
#or
plot(predyaxis~predxaxis, Dataset)
#labels: ylab="", xlab=""
```

*Plotting Line in Scatterplot*
```{r abline & points}
#for a model
abline(fit, lwd=3, col="red")
#for a known line
abline(a, b) #{intercept a, slope b}

#for non-linear models:
points(Y, fitted(fit), col="red", pch=20)

#legend for multiple lines on plot
abline()
abline()
legend(-1, legend = c("model 1", "model 2"), col=2:3, lwd=3)

```

*Boxplot*
```{r Boxplot}
plot(dataset$quantitativepredictor, dataset$anypredictor)
#Automatically creates a boxplot if first predictor is quantitative
```

*Histogram*
```{r Histogram}
hist(dataset$predictor, breaks = 10)

#with conditions
hist(Dataset$predictor[Dataset$predictor>30])
```

*Plotting for outliers*
```{r plotting for outliers}
plot(predict(fit), rstudent(fit))
```

### Simulating data
```{r random data}
set.seed(1)
x=rnorm(100)
y=2*rnorm(100)
```

### Linear Regression

```{r fitting linear models}
fit=lm(Y~x1+x2+x3+x4, data=dataset)

#no intercept
fit=lm(y~x+0)

#all predictors
fit=lm(Y~., dataset)

#all except
fit=lm(Y~.-x1-x2, dataset)

#Interactions
#between x1 & x2:
fit=lm(Y~x1*x2, dataset)

#all predictors + interaction
fit=lm(Y~.+x1:x2+x3:x4, dataset)

#QUAD/POLYNOMIAL
#Squared:
fit=lm(Y~x+I((x2)^2), dataset)
#Higher degree
fit=lm(Y~poly(x, 4))
```

*Extracting info from fits*
```{r fit info}
names(fit)
coef(fit)
summary(fit)$r.squared #R^2 of the model
summary(fit)$sigma #gives the RSE of the model
summary(fit)$coefficients[,"Pr(>|t|)"] #p-value for each coef
```

*Correlations*
```{r correlation matrix}
cor(dataset)
#NEED to exclude qualitative variable + AS FACTORS
cor(subset(dataset, select=-qualpred))
cor(Dataset[,-9]) #removing a column
```

*Function that fits & Plots at once*
```{r function to fit & plot}
regplot=function(y,x,...) #... means we have unnamed arguments but we are allowed to add more arguuments
{
  fit=lm(y~x)
  plot(x,y,...) #whatever extra parameters I add when calling the function will be added to plot()
  abline(fit, col="red")
}
regplot(Y, X, xlab="Response name", ylab="Predictor name", col="blue", pch=20)
```

*Dummy Variables*
```{r dummy variables - constrasts}
contrasts(dataset$quantpredictor)
```

*Comparing fits, ANOVA*
```{r comparing fits ANOVA}
anova(fit1, fit2)
 #large f-stat, small p-val, second model better
```

*Predictions, Confidence Intervals*
```{r predictions and confint}
#general confidence interval (95%)
confint(fit)

#predict for specific values of predictors
predict(fit, data.frame(predictor=(c(5,10,15))), interval="confidence")
predict(fit, data.frame(predictor=(c(5,10,15))), interval="prediction") 
```

*Diagnostic Plots for fit*
```{r fit diagnostic plot}
par(mfrow=c(2,2))
plot(fit)
```

### Variable Selection
```{r forward-backward-best}
#All Subset Regression
ols_step_all_possible(fit)

# Best Subset Regression
ols_step_best_subset(fit)

# Stepwise Forward Regression
ols_step_forward_p(fit)

# Stepwise Backward Regression
ols_step_backward_p(fit)
```

### Logistic Regression

*Training Set*
```{r training and test set}
#Splitting the Data Arbitrarily
train=(predictor<2005) #condition #boolean vector
test=Dataset[!train,] #subset of dataset that is "false" in train
dim(test)
Ytestval=Y[!train] #actual values for the response (category) in the test set
```

*Fitting the model*
```{r fitting model}
logist_reg_fit = glm(Y~X1+X2+X3..., data=dataset, family=binomial, subset=train)

#coefficients:
coef(logist_reg_fit)
summary(logist_reg_fit)$coef

#p-values:
summary(logist_reg_fit)$coef[ ,4]
```

*Predicting*
```{r predicting}
glm_probs = predict(logist_reg_fit, test, type="response") #vector of probabilities for each observation
#predicting over the test set using the training set model

contrasts(Y) #down vs up
#the category with value=1 is the one for which we are predicting the probability (alphabetical by default)

#no: contrast 0
#glm_pred=rep("no", (number_of_observations))
#glm_pred[glm_probs>.5]="yes" #yes constrast 1
#or
glm_pred=ifelse(glm_probs>0.5, "yes", "no") #no need to specify the number of observations

#Confusion Matrix:
table(glm_pred, Ytestval)

#Accuracy: (TP+TN)/Total
mean(glm_pred==Ytestval)
#Test set error: 
mean(glm_pred!=Ytestval)

#FNR=(FN/total positives)x100
#FPR=(FP/total negatives)x100 = 1-specificity = Type I error
#overall error=(FP+FN)/total x100
#sensitivity = TPR = (TP/Total positives) x100

#PREDICTING FOR GIVEN VALUES
predict(logist_reg_fit, newdata = data.frame(predictor1=c(1.2,1.5), predictor2=c(1.1,-0.8)), type="response")
```

### Linear Discriminant Analysis (LDA)

```{r fitting the model}
lda_fit = lda(Y~x1+x2, data=Dataset, subset=train)
lda_fit
plot(lda_fit) #produces plots of the linear discriminants
```

```{r predicting}
lda_pred = predict(lda_fit, test)
names(lda_pred) #the names are  class, contains LDA’s predictions about the movement of the market. posterior, matrix whose kth column contains the posterior probability that the corresponding observation belongs to the kth class, x contains the linear discriminants

lda_class = lda_pred$class #the class predictions of our fitted model on test data
table(lda_class, Ytestval) #predictions vs real values of test data
contrasts(Y) #the one with value=1 is the one for which we are getting the posterior probability

#Accuracy
mean(lda_class==Ytestval) #this is the accuracy
#FNR=(FN/total positives)x100
#FPR=(FP/total negatives)x100 = 1-specificity = Type I error
#overall error=(FP+FN)/total x100
#sensitivity = TPR = (TP/Total positives) x100

sum(lda_pred$posterior[,1]>=0.5) #observations with predicted proba >0.5
sum(lda_pred$posterior[,1]<0.5)

#MY OWN POSTERIOR PROBABILITY THRESHOLD
sum(lda_pred$posterior[,1]>0.9) #observations with predicted proba >0.9
my_threshlold_classes=ifelse(lda_pred$posterior[,1]>0.9, "yes", "no") #yes being the one with contrast=1
```

### Quadratic Discriminant Analysis (QDA)

```{r fitting the model}
qda_fit=qda(Y~x1+x2, data=dataset, subset = train)
qda_fit

#CV????

qda_class=predict(qda_fit, test)$class
table(qda_class, Ytestval)

#Accuracy
mean(qda_class==Ytestval)
#FNR=(FN/total positives)x100
#FPR=(FP/total negatives)x100 = 1-specificity = Type I error
#overall error=(FP+FN)/total x100
#sensitivity = TPR = (TP/Total positives) x100
```

### KNN 

```{r fitting the model}
#1.trainX:  A matrix containing the predictors associated with the training data, labeled 
#2.testX: A matrix containing the predictors associated with the data for which we wish to make predictions
#3. trainDirection: A vector containing the class labels for the training observations
#4. A value for K, the number of nearest neighbors to be used by the classifier.

#test=1:1000
#train=Dataset[-test,]
library(class)
trainX=cbind(pred1, pred2)[train,]
testX=cbind(pred1, pred2)[!train,] #from the test data
#trainX=as.matrix((Lag2)[train]) #for single predictors
#testX=as.matrix((Lag2)[!train])
trainClass=Y[train]

set.seed(1)
knn_pred=knn(trainX, testX, trainClass, k=1)
table(knn_pred, Ytestval)

mean(knn_pred==Ytestval)
#FNR=(FN/total positives)x100
#FPR=(FP/total negatives)x100 = 1-specificity = Type I error
#overall error=(FP+FN)/total x100
#sensitivity = TPR = (TP/Total positives) x100
```

